Americans who are wealthier and perform more strength training are less likely to be obese

### 2024-04-19 paper/paper.qmd
fixed pdf issue

---
title: "Americans who are wealthier and perform more strength training are less likely to be obese"
subtitle: "An analysis of how exercise habits and wealth correlate with BMI"
author: 
  - Victor Ma
thanks: "Code and data are available at: https://github.com/bestmustard/activity-bmi"
date: 03/16/2024
date-format: long
abstract: "This study examines the correlation between an individual's weekly physical activity, their income level, and their probability of being obese. The only correlation found between physical activity was with individuals who strength train having lower rates of obesity, and higher income levels tended to trend with lower obesity. The findings from this study can help researchers understand the mechanisms that drive weight loss, and also shows individuals that nutrition is a greater factor than exercise with respect to body composition."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72

@@

This analysis is conducted using R, using several R packages to facilitate my analysis and presentation. This includes tidyverse for data manipulation and visualization, knitr for report generation, modelsummary for model interpretation, and rstanarm for Bayesian regression modeling [@citeR; @Knitr; @Modelsummary; @Rstanarm; @Tidybayes; @Dplyr]. Some portions including ggplot graphs and the "Data" and "Discussion" sections were written with the help of ChatGPT4 @chatgpt.

@@

After creating the model, we can examine the summary in @fig-summary:

```{r out.height="20%"}
#| label: fig-summary
#| fig-cap: "Model summary"
#| echo: false
#| warning: false
#| message: false
obesity_likelihood <- readRDS(file ="../models/obesity_likelihood.rds")

modelsummary(obesity_likelihood)
```

#### Summary Interpretation

@@

(Intercept) (5.329): When all other variables are at their baseline levels (no physical activity and the lowest income category), the log odds of being obese are 5.329.

activityCardio + Strength (-8.164): Engaging in both cardio and strength activities is associated with an 8.164 unit decrease in the log odds of being obese compared to the baseline of no activity.

activityDouble Cardio (-4.518): Doing 300 minutes of moderate cardio or 150 minutes of vigorous cardio results in a 4.518 unit decrease in the log odds of being obese compared to the baseline.

activityNo Activity (-5.269): Having no leisure-time physical activity is associated with a 5.269 unit decrease in the log odds of being obese compared to the baseline. This negative value is unexpected, as we would typically predict no activity to increase the odds of obesity unless the baseline includes negative health behaviors even more associated with obesity.

activityStrength (-5.092): Engaging only in strength training activities is associated with a 5.092 unit decrease in the log odds of being obese compared to the baseline.

income15-25 (0.096): Being in the \$15,000 to $25,000 income bracket is associated with a 0.096 unit increase in the log odds of being obese compared to the reference income category.

income25-35 (-0.719): Being in the \$25,000 to $35,000 income bracket is associated with a 0.719 unit decrease in the log odds of being obese.

income35-50 (-0.990): Being in the \$35,000 to $50,000 income bracket is associated with a 0.990 unit decrease in the log odds of being obese.

income50-75 (-1.063): Being in the \$50,000 to $75,000 income bracket is associated with a 1.063 unit decrease in the log odds of being obese.

income75+ (-0.401): Being in the over $75,000 income bracket is associated with a 0.401 unit decrease in the log odds of being obese.

>>>>

---
title: "Americans who are wealthier and perform more strength training are less likely to be obese"
subtitle: "An analysis of how exercise habits and wealth correlate with BMI"
author: 
  - Victor Ma
thanks: "Code and data are available at: https://github.com/bestmustard/activity-bmi"
date: 04/19/2024
date-format: long
abstract: "This study examines the correlation between an individual's weekly physical activity, their income level, and their probability of being obese. The only correlation found between physical activity was with individuals who strength train having lower rates of obesity, and higher income levels tended to trend with lower obesity. The findings from this study can help researchers understand the mechanisms that drive weight loss, and also shows individuals that nutrition is a greater factor than exercise with respect to body composition."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72

@@

This analysis is conducted using R, using several R packages to facilitate my analysis and presentation. This includes tidyverse for data manipulation and visualization, knitr for report generation, modelsummary for model interpretation, and rstanarm for Bayesian regression modeling [@citeR; @Knitr; @Modelsummary; @Rstanarm; @Tidybayes; @Dplyr]. Some portions including ggplot graphs and the "Data", "Summary Interpretation", and "Discussion" sections were written with the help of ChatGPT4 @chatgpt.

@@

We can see the model summary in  @tbl-modelresults:

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = "../models/obesity_likelihood.rds")
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory model of obesity based on activity and income level"
#| warning: false

modelsummary::modelsummary(
  list(
    "Obesity Likelihood" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```
#### Summary Interpretation

@@

(Intercept) (0.06): This is the log-odds of being obese when all other variables are zero. Since it's close to zero, it suggests that when no activity is accounted for and income is at its baseline (likely the lowest income category), the likelihood of being obese is near the 50% mark on the probability scale.

activityStrength (0.18): Engaging in strength training activities is associated with a 0.18 increase in the log-odds of being obese compared to no physical activity. The positive coefficient suggests a slight association with higher obesity odds, which might be unexpected.

activityCardio (5.37): Engaging in cardio activities is associated with a substantial increase (5.37) in the log-odds of being obese. This result is counterintuitive as cardio exercises typically correlate with weight loss or lower obesity rates.

activityCardio + Strength (-2.90): Engaging in both cardio and strength activities is associated with a decrease in the log-odds of being obese. This suggests that combined exercise types might be effective at reducing the likelihood of obesity.

activityDouble Cardio (0.75): Performing a double amount of cardio is associated with an increase in the log-odds of being obese. Like the cardio variable, this positive association is counter to typical expectations.

For the income variables (reference being likely the lowest income category):

income15-25 (0.10): Being in the $15,000 to $25,000 income bracket is associated with a slight increase in the log-odds of being obese.

income25-35 (-0.72): This income bracket is associated with a decrease in the log-odds of being obese, suggesting that a higher income level correlates with lower obesity odds.

income35-50 (-0.98): Similarly, this indicates a greater decrease in the log-odds of being obese, reinforcing the trend that higher income brackets correlate with lower obesity odds.

income50-75 (-1.06): This bracket sees an even further decrease in the log-odds of obesity.

income75+ (-0.40): This suggests a smaller decrease in the log-odds of being obese for the highest income bracket compared to some of the lower brackets.
